Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-10 cordillera-cycle.tex
Develop first part of the discussion.

Independently of the choice of palaeo-temperature record used as forcing
time-series for the ice sheet model, our simulations consistently produce two
glaciation maxima during MIS~4 (61.9--55.4\,\unit{kyr}) and MIS~2
(29.5--16.9\,\unit{kyr}). These events broadly correspond in timing to the
Gladstone (MIS~4) and McConell (MIS~2) glaciation documented by geological
evidence in the northern parts of the model domain, and to the Fraser (MIS~2)
glaciation in the southern part.

However, the timing of the modelled glacial maxima strongly depends on the
choice of palaeo-temperature record used, and a more precise agreement with
timing reconstructed from geological evidence is only attained by some of the
simulations. According to radiocarbon dating, the last glacial maximum advance
of the southern Cordilleran ice sheet margin has been constrained between
00.0\,\unit{kyr} and 00.0\,\unit{kyr}. These dates have been confirmed by the
offshore sediment record, which shows an increase of xxx
betweem 00.0\,\unit{kyr} and 00.0\,\unit{kyr}.

Among simulations presented here, only those forced by the GRIP, EPICA and
Vostok records place the last glacial maximum within this range
(Fig.~\ref{fig:timeseries}, lower panel). Simulations driven by the NGRIP,
ODP~1012 and ODP~1020 records, on the opposite, produce a last glacial maximum
that pre-dates reconstructed ages by several thousand years. Concerning
simulations driven by oceanic (ODP~1012, ODP~1020) records, this early
deglaciation is caused by an early warming in the palaeo-temperature records
(Fig.~\ref{fig:timeseries}, upper panel). However, it has been shown that this
early warming is a local effect, corresponding to a weakening of the California
current \aref. The California currents is a southwards current driving cold
waters along the western coast of USA. It is thought to have weakened in the
presence of the Cordilleran ice sheet, explaining paradoxically warmer
sea-surface temperatures at the location of the ODP~1012 and ODP~1020 sites
during the last glacial maximum \aref.

This precocious glacial maximum illustrates
the complexity of climate-ice-sheet feedbacks on a regional scale, and shows
that, although closely located to the modelling domain, the ODP~1012 and
ODP~1020 proxy records records can't be reliably used as a temperature forcing
to model the last Cordilleran ice sheet.

>>>>

Independently of the choice of palaeo-temperature record used as forcing
time-series for the ice sheet model, our simulations consistently produce two
glaciation maxima during MIS~4 (61.9--55.4\,\unit{kyr}) and MIS~2
(29.5--16.9\,\unit{kyr}). These events broadly correspond in timing to the
Gladstone (MIS~4) and McConell (MIS~2) glaciation documented by geological
evidence in the northern part of the model domain, and to the Fraser (MIS~2)
glaciation in the southern part.

However, the timing of the modelled glacial maxima strongly depends on the
choice of palaeo-temperature record used, and a more precise agreement with
timing reconstructed from geological evidence is only attained in some of our
simulations. In the Puget lowland, the last glacial maximum advance of the
southern Cordilleran ice sheet margin has been constrained between by
radiocarbon dating between 17.4 and 16.4\,\unit{\chem{^{14}C}\,cal\,kyr\,BP}.
These dates have been confirmed by the offshore sediment record, which shows an
increase of glaciomarine sedimentation between 19.5 and
16.2\,\unit{\chem{^{14}C}\,cal\,kyr\,BP}.

Among simulations presented here, only those forced by the GRIP, EPICA and
Vostok records place the last glacial maximum within this range
(Fig.~\ref{fig:timeseries}, lower panel). Simulations driven by the NGRIP,
ODP~1012 and ODP~1020 records, on the opposite, produce a last glacial maximum
that pre-dates reconstructed ages by several thousand years. Concerning
simulations driven by oceanic (ODP~1012, ODP~1020) records, this early
deglaciation is caused by an early warming in the palaeo-temperature records
(Fig.~\ref{fig:timeseries}, upper panel). However, it has been shown that this
early warming is a local effect, corresponding to a weakening of the California
current \aref. The California current, driving cold waters southwards
along the western coast of North America, is thought to have weakened in the
presence of the Cordilleran ice sheet, resulting in paradoxically warmer
sea-surface temperatures at the location of the ODP~1012 and ODP~1020 sites
during the last glacial maximum \aref.

Because most of the Cordilleran ice sheet marine front lies in a different
oceanic circulation gyre than that holding the California current, it is
unlikely that it was affected by this early warming. However, the above paradox
illustrates the complexity of ice-sheet feedbacks on regional climate, and
demonstrates that, although located in the neighbourhood of the modelling
domain, the ODP~1012 and ODP~1020 palaeo-temperature proxy records can not be
used as a realistic forcing time-series to model the Cordilleran ice sheet
through the last glacial cycle.

These results makes us believe that a palaeo-temperature proxy record located
in the direct vicinity of the former Cordilleran ice sheet would be of great
help to understand its glacial history. Because, to our knowledge, no such
record yet exists, we focus the rest of our analysis on simulations forced by
temperature records from one Greenland (GRIP) and one Antarctic (EPICA) ice
cores.
